Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-08-24 demystify.tex
updates

\section{CV/Resume}
\sectioninfo{Highlight and summarize major achievements, e.g., Publications, Programming Competition Awards, and Teaching Experience.}

@@

\chapterinfo{Celebrate! Now it is your turn to evaluate the school!
Attend \emph{Open House} to learn more about the place and \emph{interview} profs---they would be much more willing to talk to you now.}

@@

You can consider applying to MS programs, which are typically easier to get in; but you likely need to pay.  Sometimes, the university that rejected you for PhD might accept you for MS.
If you get into an MS program at a school of your choice, you can contact professors to work with them. If you do well, you can ask the professor to support you to convert to PhD.

\paragraph{Should I ask for feedback?}
No, don't bother.  You will likely not get any useful feedback.  We are not willing (somtimes allowed) to reveal your evaluation results or give you feedback on how to improve your profile. \emph{So just move on}.  If you really want advice, ask your professors, collaborators, ref writers, or those who have previously applied.

@@

\chapterinfo{TAs, RAs, and fellowships are main funding sources for PhDs.  TAs are provided by the department to help with classes (e.g., grading). RAs are given by profs. to help with their research.  Fellowships, provided by the university, department, or external sources such as government or industry, give move flexibility but can be very competitive.}

@@

\chapterinfo{Not every professor, even those in CS, can advise or graduate CS PhD students.}

@@

\section{Choosing a University}\label{sec:choosing-university}

@@

\section{Choosing an Adviser}
There is no one-size-fits-all answer to this question. The best adviser is the one that you can work well with.  But how do you find such a person?

@@

\section{Why did you get rejected?}\label{sec:why-rejected}

\section{Timeline: Why are rejection letters sent so late?}\label{sec:late-rejection}
\sectioninfo{Why rejection letters are sent so late? Grad programs often wait for the accepted students to make their decisions, typically by April 15, before sending out rejection letters.}

@@

However, note that the prof. might not have the funding to buy you things. More importantly, they are not obligated to do so. It might also a need-based decision, e.g., if you need a powerful computer for your research, then the prof. might buy it for you (the lab might already has powerful servers and you can just use them for your work).  But if you just want a new computer because others have it, then the prof. might not buy it for you.

Keep in mind that these computers would be university property, which might be monitored and have certain restrictions, e.g., do not install illegal software on them (\autoref{sec:illegal-software}).  You might also have to return them when you graduate.

\section{Negotiating PhD offer (e.g., having multiple offers)?}\label{sec:negotiate}

@@

\chapter{Rankings of CS PhD programs}\label{sec:ranking}

>>>>

\section{CV/Resume}
\sectioninfo{Highlight and summarize major achievements, e.g., Publications, Programming Awards}

@@

\chapterinfo{Congrats! Now it is your turn to evaluate the school!
Attend \emph{Open House} to learn more about the place and \emph{interview} profs---they would be much more willing to talk to you now.}

@@

You can consider applying to MS programs, which are typically easier to get in; but you likely need to pay.  Sometimes, the university that rejected you for PhD might accept you for MS.
If you get into an MS program at a school of your choice, you can contact professors to work with them. If you do well, you can ask the professor to support you to convert to PhD. You will likely need to apply again, but you will have a much better chance because you have the direct support of a professor there.

\subsection{Why did you get rejected?}\label{sec:why-rejected}

\subsection{Timeline: Why are rejection letters sent so late?}\label{sec:late-rejection}
\sectioninfo{Why rejection letters are sent so late? Grad programs often wait for the accepted students to make their decisions, typically by April 15, before sending out rejection letters.}

\subsubsection{Should I ask for feedback?}
No, don't bother.  You will likely not get any useful feedback.  We are not willing (sometimes not allowed) to reveal your evaluation results or give you feedback on how to improve your profile. \emph{So just move on}.  If you really want advice, ask your professors, collaborators, ref writers, or those who have previously applied.

@@

\chapterinfo{TAs, RAs, and fellowships are main funding sources for PhDs.  TAs are provided by the department to help with classes. RAs are given by profs. to help with their research.  Fellowships, provided by the university, department, or external sources such as government or industry, give move flexibility but can be very competitive.}

@@

\chapterinfo{Not every university has PhD progarms in CS. Not every professor, even those in CS, can advise or graduate CS PhD students.}

@@

\section{Choosing a University}\label{sec:choosing-university}
\sectioninfo{Select schools based on their CS program and faculty research interests.}

@@

\section{Choosing an Adviser}
\sectioninfo{The best adviser is the one that you can work well with. But you do not know that until you start working with them.}

There is no one-size-fits-all answer to this question. The best adviser is the one that you can work well with.  But how do you find such a person?

@@

@@

However, note this is a privilege and a courtesy and not a right.  This means that the prof. is not obligated to buy you things, and that you should not expect these are automatically provided.  Also, while the prof. might be able to do so in the past, they might not be able to do so now due to budget constraints.  It might also a need-based decision, e.g., if you need a powerful computer for your research, then the prof. might buy it for you (the lab might already has powerful servers and you can just use them for your work). Finally, the prof. might have a policy on what they can buy for you, e.g., they prefer to get you desktop instead of a laptop (to encourage you to come to the office and work).

Keep in mind that these computers and equipment would be university property, which might be monitored and have certain restrictions, e.g., do not install illegal software on them (\autoref{sec:illegal-software}).  You will likely need to return them when you graduate.

\section{Negotiating PhD offer (e.g., having multiple offers)?}\label{sec:negotiate}
\sectioninfo{You will not be able to negotiate stipend, but you can ask for specific start date, TA assignment, and conference travel budget.}

@@

\chapter{CSRankings: Rankings of CS PhD programs}\label{sec:ranking}
\chapterinfo{CSRankings.org is a ranking system based on faculty publications at top CS conferences.}
